---
title: "ESM 244 Lab 1"
author: "Mukta Kelkar"
date: "1/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Read in SF Trees Data

```{r}
sf_trees <- read.csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Basic wrangling reminders

refresh some skills for data wrangling and & summary statistics using functions in the `dplyr` package

Find the top 5 highest observations of trees by legal_status, do some wrangling, make a graph.

```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>%    
  rename(tree_count = n) %>%   #new name = old name
  relocate(tree_count) %>%      #moves tree_count to the first column position. can do more variations on relocating
  slice_max(tree_count, n = 5)
```

Make a graph of the top 5 observations by legal status

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) + 
  geom_col() +
  labs(x = "Legal Status",
       y = "Tree Count") +
  coord_flip() +
  theme_minimal()
```

### A few more data wrangling refresher examples

Only keep observations for blackwood acacia trees
```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)

ggplot(data = blackwood_acacia, aes(x = longitude, y = latitude)) +
  geom_point()
```

### tidyr::separate() and unite()

Useful for combining or separating columns

```{r}
sf_trees_species <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")
```

Example of tidyr::unite()

```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool!_")
```

### Make some actual maps of blackwood acacia trees in SF

`st_as_sf()` to convert latitude & longitude to spatial coordinates.

```{r}
blackwood_acacia_sp <- blackwood_acacia %>% 
  drop_na(longitude, latitude) %>% 
  st_as_sf(coords = c("longitude", "latitude"))

st_crs(blackwood_acacia_sp) = 4326

ggplot(data = blackwood_acacia_sp) +
  geom_sf(color = "darkgreen")
```

Read in SF roads shpaefile:

```{r}
sf_map <- read_sf(here("data", "sf_map", "tl_2017_06075_roads.shp"))

st_transform(sf_map, 4326)

ggplot(data = sf_map) +
  geom_sf()
```

Combine blackwood acacia tree observations & SF roads map:

```{r}
ggplot() +
  geom_sf(data = sf_map, size = 0.1, color = "darkgray") +
  geom_sf(data = blackwood_acacia_sp, color = "green", size = 0.5) +
  theme_void()
```

Now an interactive map

```{r}
tmap_mode("view")

tm_shape(blackwood_acacia_sp) + 
  tm_dots()
```

